Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2023-07-30 ba_resubmit/response_to_reviewers3.tex

\begin{document}

	% ****************** TITLE ****************************************

	\title{Response to Reviewer Comments}
	\maketitle

We thank the Editor and Associate Editor for their assessment and constructive feedback.

@@

%Thank you for the thoughtful feedback. We have made several broad changes to address the Editor's and Associate Editor's overall critiques, and have addressed each specific comment from the reviewer.

\section{AE Comments}

\pointRaised{AE}{I must add that reading the paper still feels like a chore, in part because the notation is so heavy, which may be hard to avoid, but also, in part, because some wording choices make certain sentences hard to parse, which can certainly be improved.}

\reply{We have made several revisions to the paper, as described in our responses below. 
}

@@

``Consider two data files $A$ and $B$ comprising $n_A$ and $n_B$ records, respectively, and including $F$ linkage variables measured in both files. For $i=1, \dots, n_A$, let record $i$ be given by $A_i = (A_{i1}, \dots, A_{iF})$, so that $A = (A_i : i = 1, \dots, n_A)$.  Similarly, for $j=1, \dots, n_B$, let record $j$ be given by $B_j = (B_{j1}, \dots, B_{jF})$, so that $B = (B_j : j = 1, \dots, n_B)$.  Without loss of generality, denote files such that $n_A \geq n_B$.". See Section 2. 
}

\setcounter{responsectr}{0}
\section{Associate Editor Comments}

	\pointRaised{R}{%
	For the missing data treatment now included just before Section 2.1, is the assumption truly missing at random, or missing completely at random? Or, does this distinction not matter because of the independence assumed across elements of the comparison vector?}

	\reply{
		%Because we assume that the agreement level for each field is conditionally independent given the matching status, MAR and MCAR are equivalent here. For simplicity, we have chosen to use MCAR in the article. 
		We have included the following explanation of missingness in Section 2.

@@

%\reply{No, in \text{fabl}, we do not integrate out $\pi$ as done in \cite{sadinle_bayesian_2017} because this leads to a sequential sampler. 
%
%(1) Yes, our alternative approach is different than directly integrating over the prior distribution $\pi.$ 
%%We have attempted to make this more clear in our revised version. 
%In short, our prior distribution over $\bm{Z}$ leads to a Gibbs sampler that leads to parallel updates, which contrasts that of \cite{sadinle_bayesian_2017}. Of course, we could integrate out $\pi$ in our approach, however, this leads to sequential updates, which would be slow in practice. See Appendix A, page 27. 
%
%(2) Yes, our alternative approach is justified. We provide a justification in Appendix A, page 27 regarding why the updates are independent and not sequential. Thank you for the excellent questions, which have greatly improved our paper. 
%
%}

\pointRaised{R}{%
	I do not understand the statement just below Equation (10): “When j has no match in $X_1$, we write $(n_1 +j, j) \in h_{P+1}$” My understanding of these patterns is that they are based on observed comparison vectors without consideration of Z (matches). In the second paragraph of Section 4.3, the H notation includes the matches (Z), in notation and definition that seems to conflict with the statement just below Equation (10). This also comes into play in Equation (16).}

	%% RCS: Tried to answer this from the original paper and the updated paper. Tried to make it shorter and close it out very quickly.

	\reply{Yes, the comparison vectors are created (and therefore, patterns are assigned) without regard to $\bm{Z}$. We have removed that sentence, and updated the equations in the Section 4.3 to avoid confusion.}

@@

\pointRaised{R}{%
	Third paragraph of Section 4.1: The authors claim they are computing “sufficient statistics”. What exactly are these statistics sufficient for?
	}

	\reply{We have revised the text to state ``summary statistics" instead of ``sufficient statistics."

		%We show in the revised Section 4.2 that we can write conditional likelihoods  and posterior updates for $\bm{m}$, $\bm{u}$, and $\bm{Z}$ using the statistics in $\tilde{\gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{N} \}$. However, we cannot express the full likelihood in (14) through these statistics. Therefore, we have revised to text to say ``summary statistics" instead of  ``sufficient statistics."

@@

	%Thank for pointing out our typo. We have revised the text to state ``summary statistics" instead of ``sufficient statistics."
	}

\pointRaised{R}{%
	First paragraph of Section 4.2: The authors state: “Posterior calculations still attribute the appropriate weight to all records through the summary statistics…” What is meant by the term “weight”? Which records are appropriately weighted – those in X2?}

		%% Comment: Brian, you have not answered the reviewer's question regarding what is meant by weight and what records are appropriately weighted. \textcolor{red}{Please revise to directly answer the question in the revised text with equation numbers. To my understanding, the weights have been removed and section 4 has been rewritten }}
%	\reply{
%	\textcolor{teal}{In the original manuscript, we consider all records in the first file ($X_{1}$) that share an agreement pattern with the second data file ($X_{2}$). In this situation, all these records have the same Fellegi-Sunter likelihood ratio weight, $w_{ij}$. Given this result, we denoted the weight $w_p$ since these records map to the same agreement pattern.} 
%	
%	\textcolor{teal}{In the revised manuscript, we define $m_p$ and $u_p,$ which are the 
%	 probabilities that records $A_i$ and $B_j$ form agreement pattern $p$ given that
% they are a match and non-match, respectively. For each pattern $p,$ we define $w_p = m_p/u_p.$ We hope that this will be more clear than our previous presentation of the material. See Section 4.1, pages 8--9.}
% }
% 
	\reply{Our intended meaning was ``All record pairs contribute to the likelihood through the summary statistics."

	In our revised submission, we have re-expressed the likelihood from (6a) through the perspective of agreement patterns in (14). This form of the likelihood is then used to express contributions to the likelihood for the $\bm{m}$ and $\bm{u}$ parameters in (15) and for $\bm{Z}$ in (17), both of which use the summary statistics. We hope that this revision makes this point more clear.}

@@

	\reply{

%	In the revised paper, ``vector" is replaced with ``matrix" when referring to $\gamma.$ (See R.13 regarding making the paper consistent below).

	In the original manuscript, ``those comparison vectors" refer to removing the larger (and expensive) $\Gamma^{ab}$ from memory and continuing our calculations with the compressed comparison vectors $\tilde{\Gamma}^{ab}.$

	The revised text states: ``Then, we conduct hashing, obtain the compressed comparison matrix, $\tilde{\gamma}^{ab}$, and remove the memory-intensive comparison matrix, $\gamma^{ab}$, before continuing with the next batch of data. See Section 4.3.}

@@

%	 \textcolor{red}{Question for Brian: is lowercase $\gamma$ always a matrix? Let's make sure we're not swapping between vectors and matrices. In the previous version this was a vector and now its listed as a matrix. I would argue that a matrix should use capital notation, such as $\Gamma.$}

%\reply{%
%	We have revised that sentence to read ''We then conduct hashing, obtain the compressed $\tilde{\Gamma}^{ab}$ for later calculations, and delete the larger $\Gamma^{ab}$ from memory before continuing with the next chunk of data."
%}

\pointRaised{R}{%
	Where does $R^{SEI, cd}$ come into play in the partitioned algorithm presented in Equations (13) and (14)? I recommend that the authors either refrain from suppressing the SEI notation or further explain how the SEI algorithm has changed the quantities in these equations.}

\reply{To make the SEI algorithm more clear, we have revised the ordering of Sections 4, so that we have 1) hashing, 2) posterior inference, 3) chunkwise computation of the comparison matrix, and 4) SEI. With this new ordering, and with the newly provided equations (15) through (19), we hope it is clearer that that the contribution of \emph{all} record pairs is recorded through the summary statistics in $\mathcal{N}$. In particular, (16a), (16b), and (18) depend only $\mathcal{N}$. The SEI algorithm only affects the step shown in (19). 
	}

@@

\pointRaised{R}{%
	Equation (6a): The indices do not match the subscripts in the indicator function in each summand, or their standard meaning in table 1.}

\reply{%
	The typo has been fixed.
}

\pointRaised{R}{%
	Generally, the authors seem to arbitrarily use upper and lower case z interchangeably in function definitions.}

\reply{%
	We use $\bm{Z}$ when discussing a random quantity. In all instances where we discuss the random variable $Z_j$ taking on a particular value, we now use $q$ to avoid confusion and to follow notation used by \cite{sadinle_bayesian_2017}.
}

@@

\pointRaised{R}{%
Equations (8) and (9): Should the weights have superscript (s) (as the Zs do)?}

\reply{%
	This has been fixed. See Equations (9) and (10).
}

\pointRaised{R}{%
	Gamma is in some places described as a set and in others as a matrix (particularly in sections 4.2 and 5.1).
}

\reply{%
	The comparison matrix $\gamma$ is comprised of the comparison vectors for all record pairs in $A \times B$. All references to $\gamma$ as a set have been removed. See Section 2, where we define both the comparison vector and comparison matrix.
}

### 2023-07-30 ba_resubmit/kundinger_resubmit3.tex
edits

\documentclass[ba]{imsart}
%
\pubyear{0000}
\volume{00}
\issue{0}
\doi{0000}
%\arxiv{}
\firstpage{1}
\lastpage{1}

@@

%% *** Frontmatter *** 
\linenumbers
\begin{frontmatter}
\title{Efficient and Scalable Bipartite Matching with Fast Beta Linkage  (fabl)}
%\title{A Sample Document\thanksref{T1}}
%\thankstext{T1}{<thanks text>}
\runtitle{Efficient and Scalable Bipartite Matching with Fast Beta Linkage  (fabl)}

@@

%% ** Keywords **
\begin{keyword}%[class=MSC]
\kwd{data fusion}
\kwd{distributed computing}
\kwd{entity resolution}
\kwd{hashing}
\kwd{record linkage}
\end{keyword}

@@

Before conducting data analysis, it is often necessary to identify duplicate records across two data files. This is an increasingly important task in ``data cleaning'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others \citep{christen_2012, gutman2013bayesian, DalzellReiter18, tang2020}. In this article, we consider bipartite record linkage, which merges two data files that contain duplications across, but not within, the respective data files.

@@

Intuitively, matching records (those that refer to the same entity) should have similar values of the linking variables; records that are nonmatching should have dissimilar values. \cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\gamma_{ij}$ computed for each record pair $(i,j)$ in $A \times B.$  Specifically, they define $\gamma_{ij} = (\gamma_{ij}^1, \ldots, \gamma_{ij}^F),$ where each $\gamma_{ij}^f$ is a value indicating the similarity of field $f$ for records $A_i$ and $B_j$. We define the comparison matrix $\gamma \in \mathbb{R}^{n_A n_B F}$ as the collection of all record pairs $\gamma_{ij}$. Additionally, we let $\Gamma_{ij}$ represent a random variable for the comparison vector for $A_i$ and $B_j$, such that $\gamma_{ij}$ is a realization of $\Gamma_{ij}$.

@@

%Let $\Gamma_{ij}$ represent a random variable for the comparison vector for arbitrary $A_i$ and $B_j$.  Thus, $\gamma_{ij}$ is a realization of $\Gamma_{ij}$.

For modeling the collection of $n_An_B$ random variables $\Gamma_{ij}$, \cite{fellegi_theory_1969} employ two independence assumptions: first, that comparison vectors are conditionally independent given their matching status, and second, that the matching status of the record pairs are independent. Using these independence assumptions,  one specifies a mixture model for each $\Gamma_{ij}$ \citep[e.g., as in ][]{winkler_state_1999,jaro1989,larsen_2001,enamorado2019using}.  We have
\begin{subequations}
\begin{align}
&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \label{eqn:fs_model} \\
&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \label{eqn:m_dist}\\
&\Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(\lambda). \label{eqn:lambda_dist}
\end{align}
\end{subequations}
Here, $\mathcal{M}$ and $\mathcal{U}$ are the distributions for matching and nonmatching record pairs, $\bm{m}$ and $\bm{u}$ are their respective sets of parameters, and $\lambda$ is the marginal probability that a record pair is a match. When using comparison vectors with discrete agreement levels, $\mathcal{M}$ and $\mathcal{U}$ are collections of independent multinomial distributions for each linkage feature. Accordingly, $\bm{m} = (\bm{m}_1, \ldots, \bm{m}_F)$, where $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$ and $m_{fl} = p(\gamma_{ij}^f = l|\Delta_{ij} = 1)$ for all fields $f$ and agreement levels $l$. The $\bm{u}$ parameters are defined similarly, with $u_{fl} = p(\gamma_{ij}^f = l|\Delta_{ij} = 0)$.

@@

Instead of the prior over $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we follow \cite{wortman2019} and use independent priors for each component $Z_j$. However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$, we use the fast Beta linkage (\texttt{fabl}) prior below. For each $Z_j$, we have
%\begin{subequations}
\begin{align}
	p(Z_j = q| \pi) &= \begin{cases} 
	\frac{1}{n_A}\pi,  & q \leq n_A;\\
	1-\pi, &  q  = n_A + j;  \\
\end{cases} \label{eqn:fast_beta_prior}  \\
\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}) \nonumber
%\label{eqn:fast_beta_prior2}.
\end{align}
%\end{subequations}
We can interpret (\ref{eqn:fast_beta_prior}) as follows: record $B_j$ has some match in $A$ with probability $\pi$, and each record $A_i$ is equally likely to be that match. The hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about  the proportion of records in $B$ that have matches in $A.$

In the \cite{wortman2019} flat prior, each value $\{1, \ldots, n_A, n_A +j\}$ is a priori equally likely for $Z_j$. This amounts to a prior probability of $n_A / (n_A + 1)$ that record $B_j$ has a match in $A$. In our preliminary studies, this highly informative prior weighting on matching can result in overly high match rates. Hence, we prefer (\ref{eqn:fast_beta_prior}). We also note that the flat prior is equivalent to a special case of the fast Beta prior with $\pi$ fixed at the mean of a $\text{Beta}\left(1, 1 / n_A \right)$ random variable.

Linkage with \texttt{fabl} is conducted at the record level, rather than at the record pair level, as in the \cite{fellegi_theory_1969} model. That is, $\pi$  in (\ref{eqn:fast_beta_prior}) under \texttt{fabl} estimates the proportion of records in $B$ that have matches, whereas $\lambda$ in (\ref{eqn:lambda_dist}) in the \cite{fellegi_theory_1969} model estimates the proportion of record pairs that are matches. In the bipartite case, we conjecture that some analysts will find $\pi$ to be more a interpretable parameter than $\lambda$. In this setting, there are at most $n_B$ matching pairs out of $n_A n_B$ total pairs, meaning that $\lambda$ is bounded above by $1/n_A$ and tends towards 0 as the size of the linkage task grows. Additionally, while the \cite{fellegi_theory_1969} model makes $n_A  n_B$ independent matching decisions and \texttt{BRL} makes $n_B$ dependent matching decisions, \texttt{fabl} strikes a middle ground between the two, making $n_B$  independent matching decisions. As shown in Sections \ref{sec:simulations} and \ref{sec:case-studies}, this allows \texttt{fabl} to fit a Bayesian record linkage model like \texttt{BRL} while making computational efficiency gains possible by exploiting independence.

%\begin{table}[t!]
%	\centering
%	\begin{tabular}[t!]{ll}
%		Symbol & Description \\
%		\hline
%		$A, B$ & data files\\
%		$i \in 1, \ldots, n_A $ & index over records in $A$\\
%		$j \in 1, \ldots, n_B $ & index over records in $B$\\
%		$f \in 1, \ldots F$ & index over fields used for comparisons \\
%		$l \in 1, \ldots L_f$ & index over agreement levels for feature $f$ \\
%		$n_{AB}$ & number of matches between $A$ and $B$\\
%		$\gamma_{ij}$ & comparison vector for records $A_i$ and $B_j$ \\
%		$Z_j = i$ & records $A_i$ and $B_j$ match \\
%		$Z_j = n_A + j$ & record $B_j$ has no match in $A$ \\
%		$m_{fl}$ & $p(\gamma_{ij}^f = l | Z_j = i)$ \\
%		$u_{fl}$ & $p(\gamma_{ij}^f = l | Z_j \neq i)$ \\
%		$\pi$ & probability that a record in $B$ has a match in $A$ \\
%		\hline
%	\end{tabular}\caption{Summary of model notation.}\label{table_notation_A}
%\end{table}
%
For clarity, we present our full model for \texttt{fabl} below:
\begin{subequations}
\begin{align}
	\mathcal{L}(\bm{Z}, \bm{m}, \bm{u} \mid \gamma) &= \prod_{i=1}^{n_A}  \prod_{j=1}^{n_B}\prod_{f=1}^{F}\prod_{l=1}^{L_f}\left[  m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}, \label{eqn:likelihood}\\
	\bm{m}_f &\sim \text{Dirichlet}(\alpha_{f1}, \ldots, \alpha_{f L_f}), \forall f = 1, \ldots, F, \label{eqn:m} \\
	\bm{u}_f &\sim \text{Dirichlet}(\beta_{f1}, \ldots, \beta_{f L_f}),\forall f = 1, \ldots, F,  \label{eqn:u}\\
	p(Z_j = q| \pi)  &=
	\begin{cases} 
		\frac{1}{n_A}\pi,  & q \leq n_A; \\
		1-\pi, &  q  = n_A + j; \\
	\end{cases} \label{eqn:z}\\
	\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})\label{eqn:pi}.
\end{align}
\end{subequations}

@@

We estimate the posterior distribution of the parameters in (\ref{eqn:likelihood} - \ref{eqn:pi}) using a Gibbs sampler. Appendix \ref{app:derivations} presents derivations for the full conditional distributions. We initialize $\bm{m}$, $\bm{u}$ and $\pi$ from random draws from their prior distributions, and initialize $\bm{Z}$ to reflect no matches across data files; that is, $\bm{Z} = (n_A + 1, \ldots, n_A + n_B)$. Denote the current draw of the sampler by $(s)$ and the updated draw by $(s+1)$. To update $\bm{m}_f$ and $\bm{u}_f$ for all $f = 1, \ldots, F$, sample
\begin{subequations}
\begin{align}
	\bm{m}_f^{(s+1)}|\gamma, \bm{Z}^{(s)}, \bm{u}^{(s)}, \pi^{(s)} &\sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}^{(s)}), \ldots, \alpha_{fL_f}(\bm{Z}^{(s)})), \label{eqn:m_update} \\
	\bm{u}_f^{(s+1)}|\gamma, \bm{Z}^{(s)}, \bm{m}^{(s)}, \pi^{(s)} &\sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})), \label{eqn:u_update} \\
	\text{where }\alpha_{fl}(\bm{Z}^{(s)})&= \alpha_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} = i), \label{eqn:alpha_update} \\
	\text{and } \beta_{fl}(\bm{Z}^{(s)})&=  \beta_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B}  I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} \neq i) \label{eqn:beta_update}.
	%where $\alpha_{fl}(\bm{Z}^{(s)})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} = i)$, and $\beta_{fl}(\bm{Z}^{(s)})= \\ \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} \neq i)$.
\end{align}
\end{subequations}
Next, we sample $\pi$ from its full conditional, given by
\begin{align}
	\pi^{(s+1)}|\gamma, \bm{Z}^{(s)}, \bm{m}^{(s+1)}, \bm{u}^{(s+1)} \sim  \text{Beta}(n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi}).
\end{align}
where $n_{AB}(\bm{Z}^{(s)}) = \sum_{j=1}^{n_B} I(Z_j^{(s)} \leq n_A)$.

@@

After computing all comparison vectors within a block, the modeler can further reduce the number of comparison vectors used in the linkage algorithm through indexing. For example, one might only consider pairs with a certain similarity score on a field deemed to be important, like last name, or only pairs that exactly match on a specified number of fields. However, the impact of indexing on model parameters is not well understood; \cite{murray2016probabilistic} reviewed this issue in the context of the classical Fellegi-Sunter model, leaving the effect of indexing on Bayesian record linkage models to future work.

With \texttt{fabl}, we introduce two techniques to further expand the scalability of probabilistic record linkage. First, we propose hashing methods that allow us to compute summary statistics that reduce the computational complexity of the Gibbs sampler and the memory requirements of the storing the comparison vectors. Second, we introduce storage efficient indexing, which reduces the memory costs associated with unlikely matches. 
%For convenience, Table \ref{table_notation_B} summarizes the notation introduced throughout this section.

>>>>

\documentclass[ba]{imsart}
%
\pubyear{2021}
\volume{TBA}
\issue{TBA}
% \doi{0000}
%\arxiv{}
\firstpage{1}
\lastpage{1}

@@

%% *** Frontmatter *** 
\linenumbers
\begin{frontmatter}
\title{Efficient and Scalable Bipartite Matching with Fast Beta Linkage  (fabl)}
\runtitle{Efficient and Scalable Bipartite Matching with Fast Beta Linkage  (fabl)}

@@

%% ** Keywords **
\begin{keyword}%[class=MSC]
\kwd{data fusion}
\kwd{data cleaning}
\kwd{entity resolution}
\kwd{hashing}
\kwd{record linkage}
\end{keyword}

@@

Before conducting data analysis, it is often necessary to identify duplicate records across two data files. This is an increasingly important task in ``data cleaning'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others \citep[e.g.,][]{christen_2012, gutman2013bayesian, DalzellReiter18, tang2020}. In this article, we consider bipartite record linkage, which merges two data files that contain duplications across, but not within, the respective data files.

@@

Intuitively, matching records (those that refer to the same entity) should have similar values of the linking variables; records that are nonmatching should have dissimilar values. \cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\gamma_{ij}$ computed for each record pair $(i,j)$ in $A \times B.$  Specifically, they define $\gamma_{ij} = (\gamma_{ij}^1, \ldots, \gamma_{ij}^F),$ where each $\gamma_{ij}^f$ is a value indicating the similarity of field $f$ for records $A_i$ and $B_j$. We define the comparison matrix $\gamma \in \mathbb{R}^{n_A n_B \times F}$ as the collection of all record pairs $\gamma_{ij}$. 
%Additionally, we let $\Gamma_{ij}$ represent a random variable for the comparison vector for $A_i$ and $B_j$, such that $\gamma_{ij}$ is a realization of $\Gamma_{ij}$.

@@

Let $\Gamma_{ij}$ represent a random variable for the comparison vector for arbitrary $A_i$ and $B_j$.  Thus, $\gamma_{ij}$ is a realization of $\Gamma_{ij}$. For modeling the collection of $n_An_B$ random variables $\Gamma_{ij}$, \cite{fellegi_theory_1969} employ two independence assumptions: first, that comparison vectors are independent given the matching status of the record pair, and second, that the matching status of each record pair is independent of the matching status of other pairs. Using these independence assumptions, one specifies a mixture model for each $\Gamma_{ij}$ \citep[e.g., as in ][]{winkler_state_1999,jaro1989,larsen_2001,enamorado2019using}.  We have
\begin{subequations}
\begin{align}
&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \label{eqn:fs_model} \\
&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \label{eqn:m_dist}\\
&\Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(\lambda). \label{eqn:lambda_dist}
\end{align}
\end{subequations}
Here, $\mathcal{M}$ and $\mathcal{U}$ are the distributions for matching and nonmatching record pairs, $\bm{m}$ and $\bm{u}$ are their respective sets of parameters, and $\lambda$ is the marginal probability that a record pair is a match. When using comparison vectors with discrete agreement levels, $\mathcal{M}$ and $\mathcal{U}$ are collections of independent multinomial distributions for each linkage feature. Accordingly, $\bm{m} = (\bm{m}_1, \ldots, \bm{m}_F)$, where $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$ and $m_{fl} = p(\Gamma_{ij}^f = l|\Delta_{ij} = 1)$ for all fields $f$ and agreement levels $l$. The $\bm{u}$ parameters are defined similarly, with $u_{fl} = p(\Gamma_{ij}^f = l|\Delta_{ij} = 0)$.

@@

Instead of the prior over $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we follow \cite{wortman2019} and use independent priors for each component $Z_j$. However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$, we use the fast beta linkage (\texttt{fabl}) prior below. For each $Z_j$, we have
%\begin{subequations}
\begin{align}
	p(Z_j = q| \pi) &= \begin{cases} 
	\frac{1}{n_A}\pi,  & q \leq n_A;\\
	1-\pi, &  q  = n_A + j;  \\
\end{cases} \label{eqn:fast_beta_prior}  \\
\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}). \nonumber
%\label{eqn:fast_beta_prior2}.
\end{align}
%\end{subequations}
We can interpret (\ref{eqn:fast_beta_prior}) as follows: record $B_j$ has some match in $A$ with probability $\pi$, and each record $A_i$ is equally likely to be that match. The hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about  the proportion of records in $B$ that have matches in $A.$

In the \cite{wortman2019} flat prior, each value $\{1, \ldots, n_A, n_A +j\}$ is a priori equally likely for $Z_j$. This amounts to a prior probability of $n_A / (n_A + 1)$ that record $B_j$ has a match in $A$. In our preliminary studies, we have found that this highly informative prior weighting on matching can result in overly high match rates. Hence, we prefer (\ref{eqn:fast_beta_prior}). We also note that the flat prior is equivalent to a special case of the fast beta prior with $\pi$ fixed at the mean of a $\text{Beta}\left(1, 1 / n_A \right)$ random variable.

Linkage with \texttt{fabl} is conducted at the record level, rather than at the record pair level, as in the \cite{fellegi_theory_1969} model. That is, $\pi$  in (\ref{eqn:fast_beta_prior}) under \texttt{fabl} estimates the proportion of records in $B$ that have matches, whereas $\lambda$ in (\ref{eqn:lambda_dist}) in the \cite{fellegi_theory_1969} model estimates the proportion of record pairs that are matches. In the bipartite case, we conjecture that some analysts will find $\pi$ to be a more interpretable parameter than $\lambda$. In this setting, there are at most $n_B$ matching pairs out of $n_A n_B$ total pairs, meaning that $\lambda$ is bounded above by $1/n_A$ and tends towards 0 as the size of the linkage task grows. Additionally, while the \cite{fellegi_theory_1969} model makes $n_A  n_B$ independent matching decisions and \texttt{BRL} makes $n_B$ dependent matching decisions, \texttt{fabl} strikes a middle ground between the two, making $n_B$  independent matching decisions. As shown in Sections \ref{sec:simulations} and \ref{sec:case-studies}, this allows \texttt{fabl} to fit a Bayesian record linkage model like \texttt{BRL} while making computational efficiency gains possible by exploiting independence.

%\begin{table}[t!]
%	\centering
%	\begin{tabular}[t!]{ll}
%		Symbol & Description \\
%		\hline
%		$A, B$ & data files\\
%		$i \in 1, \ldots, n_A $ & index over records in $A$\\
%		$j \in 1, \ldots, n_B $ & index over records in $B$\\
%		$f \in 1, \ldots F$ & index over fields used for comparisons \\
%		$l \in 1, \ldots L_f$ & index over agreement levels for feature $f$ \\
%		$n_{AB}$ & number of matches between $A$ and $B$\\
%		$\gamma_{ij}$ & comparison vector for records $A_i$ and $B_j$ \\
%		$Z_j = i$ & records $A_i$ and $B_j$ match \\
%		$Z_j = n_A + j$ & record $B_j$ has no match in $A$ \\
%		$m_{fl}$ & $p(\gamma_{ij}^f = l | Z_j = i)$ \\
%		$u_{fl}$ & $p(\gamma_{ij}^f = l | Z_j \neq i)$ \\
%		$\pi$ & probability that a record in $B$ has a match in $A$ \\
%		\hline
%	\end{tabular}\caption{Summary of model notation.}\label{table_notation_A}
%\end{table}
%
For clarity, we present the full model for \texttt{fabl} below:
\begin{subequations}
\begin{align}
	\mathcal{L}(\bm{Z}, \bm{m}, \bm{u} \mid \gamma) &= \prod_{i=1}^{n_A}  \prod_{j=1}^{n_B}\prod_{f=1}^{F}\prod_{l=1}^{L_f}\left[  m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}, \label{eqn:likelihood}\\
	\bm{m}_f &\sim \text{Dirichlet}(\alpha_{f1}, \ldots, \alpha_{f L_f}), \forall f = 1, \ldots, F, \label{eqn:m} \\
	\bm{u}_f &\sim \text{Dirichlet}(\beta_{f1}, \ldots, \beta_{f L_f}),\forall f = 1, \ldots, F,  \label{eqn:u}\\
	p(Z_j = q| \pi)  &=
	\begin{cases} 
		\frac{1}{n_A}\pi,  & q \leq n_A; \\
		1-\pi, &  q  = n_A + j; \\
	\end{cases} \label{eqn:z}\\
	\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})\label{eqn:pi}.
\end{align}
\end{subequations}

@@

We estimate the posterior distribution of the parameters in (\ref{eqn:likelihood} - \ref{eqn:pi}) using a Gibbs sampler. Appendix \ref{app:derivations} presents derivations for the full conditional distributions. We initialize $\bm{m}$, $\bm{u}$ and $\pi$ from random draws from their prior distributions, and initialize $\bm{Z}$ to reflect no matches across data files; that is, $\bm{Z} = (n_A + 1, \ldots, n_A + n_B)$. Denote the current draw of the sampler by $(s)$ and the updated draw by $(s+1)$. To update $\bm{m}_f$ and $\bm{u}_f$ for all $f = 1, \ldots, F$, we sample
\begin{subequations}
\begin{align}
	\bm{m}_f^{(s+1)}|\gamma, \bm{Z}^{(s)}, \bm{u}^{(s)}, \pi^{(s)} &\sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}^{(s)}), \ldots, \alpha_{fL_f}(\bm{Z}^{(s)})), \label{eqn:m_update} \\
	\bm{u}_f^{(s+1)}|\gamma, \bm{Z}^{(s)}, \bm{m}^{(s)}, \pi^{(s)} &\sim \text{Dirichlet}(\beta_{f1}(\bm{Z}^{(s)}), \ldots, \beta_{fL_f}(\bm{Z}^{(s)})), \label{eqn:u_update} \\
	\text{where }\alpha_{fl}(\bm{Z}^{(s)})&= \alpha_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} = i), \label{eqn:alpha_update} \\
	\text{and } \beta_{fl}(\bm{Z}^{(s)})&=  \beta_{fl} + \sum_{i=1}^{n_A}\sum_{j=1}^{n_B}  I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} \neq i) \label{eqn:beta_update}.
	%where $\alpha_{fl}(\bm{Z}^{(s)})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} = i)$, and $\beta_{fl}(\bm{Z}^{(s)})= \\ \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f) I(\gamma_{ij}^f = l) I(Z_j^{(s)} \neq i)$.
\end{align}
\end{subequations}
Next, we sample $\pi$ from its full conditional, given by
\begin{align}
	\pi^{(s+1)}|\gamma, \bm{Z}^{(s)}, \bm{m}^{(s+1)}, \bm{u}^{(s+1)} \sim  \text{Beta}(n_{AB}(\bm{Z}^{(s)}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}^{(s)}) + \beta_{\pi}).
\end{align}
where $n_{AB}(\bm{Z}^{(s)}) = \sum_{j=1}^{n_B} I(Z_j^{(s)} \leq n_A)$.

@@

After computing all comparison vectors within a block, the modeler can further reduce the number of comparison vectors used in the linkage algorithm through indexing. For example, one might only consider pairs with a certain similarity score on a field deemed to be important, like last name, or only pairs that exactly match on a specified number of fields. However, the impact of indexing on model parameters is not well understood; \cite{murray2016probabilistic} reviewed this issue in the context of the classical \cite{fellegi_theory_1969} model, leaving the effect of indexing on Bayesian record linkage models to future work.

With \texttt{fabl}, we introduce two techniques to further expand the scalability of probabilistic record linkage. First, we propose hashing methods that allow us to compute summary statistics that reduce the computational complexity of the Gibbs sampler and the memory requirements of storing the comparison vectors. Second, we introduce storage efficient indexing, which reduces the memory costs associated with unlikely matches. 
%For convenience, Table \ref{table_notation_B} summarizes the notation introduced throughout this section.
